---
title: "Programming Basics"
author: "Nils Indreiten"
date: "29/03/2022"
output: html_document
---

# Object Inspection and exploration:


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Consider the following object, diamonds. We
can inspect it using str():

```{r}
str(diamonds)
```

A more readable way to achieve the same is using
glimpse():

```{r}
diamonds %>% 
  glimpse()
```

We can use str() even with models, consider the 
following:

```{r}
lm_mod <- lm(mpg~., data=mtcars)
str(lm_mod) # Increase the level of depth by assigning a number after lm_mod for example: str(lm_mod,1)
```

We can look more into the model object by doing the following:

```{r}
lm_mod$coefficients
```

We can even summarize it:

```{r}
summary(lm_mod)
```
We can even assign the summary object into an 
object and inspect that:

```{r}
lm_mod_summary <- summary(lm_mod)
str(lm_mod_summary)
```

If we pull the coefficients from this object, we 
dont only get the values but the table
thats printed in the summary. We can easily
get it ready for publishing:

```{r}
lm_mod_summary$coefficients %>% 
  kableExtra::kable(digits = 2)
```

# Methods:

The methods() function will allow us to see how
the function will behave on different objects:

```{r}
methods('summary')
```

# S4 Classes
















